```{r {{id}}-setup, echo=FALSE}
#A species is randomly sampled if this template is run interactively.
#Else, the id is equal to the {{id}} that is passed through knit_expand
if (interactive()) {
  this_id <- sample(unique(iris$Species), 1)
} else {
  this_id <- "{{id}}"
}
selection <- iris %>%
  filter(Species == this_id)
```

## Analysis of the iris dataset for the `r stringr::str_to_title(this_id)` species

The average sepal width and length are `r round(mean(selection$Sepal.Width),2)` and `r round(mean(selection$Sepal.Width),2)` respectively. 

```{r {{id}}-sepal, echo = FALSE, warning=FALSE, message =FALSE, fig.height=1, fig.width=3}
selection %>%
  ggplot(aes(x = Sepal.Width, y = Sepal.Length)) +
  geom_point()
```

```{r {{id}}-regression, echo = FALSE, warning=FALSE, message=FALSE}
reg <- lm(data = selection,
   formula = Sepal.Width ~ Sepal.Length + Petal.Length + Petal.Width)
summary_regression <- summary(reg)$coefficients
#we're not interested in the intercept
significant <- which(summary_regression[-1,4] < 0.05)

if (length(significant) == 0) {
   conclusion <- "none of the covariates are significant"
} else if (length(significant) == 1) {
     conclusion <- paste("only", names(significant), "is significant")
} else {
       conclusion <- paste(length(significant), "of the covariates are significant")
       }
summary_regression %>%
  knitr::kable(caption = "regression results",
        booktabs = TRUE,
        digits = 3) %>%
  kableExtra::kable_styling(font_size = 7)
```

A simple linear regression model that predicts the sepal width based on sepal length, petal length and petal width shows that `r conclusion` (see Table \@ref(tab:{{id}}-regression)).
