---
author: "Research Institute for Nature and Forest"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2: default
biblio-style: "apalike"
link-citations: true
params:
    species: "setosa"
---


```{r echo=FALSE}
library(tidyverse)
library(kableExtra)
library(rprojroot)
options(knitr.table.format = function() {
  if (knitr::is_latex_output())
    "latex" else "pipe"
})
this_species <- params$species
selection <- iris %>%
  dplyr::filter(Species == this_species)
#this template shows how to make a dynamic title for the document
#see `title:` below
# the dynamic bibliography is only required now since there is one references.bib file for all inbomd_examples
```
---
title: "Analysis of the iris dataset for the `r stringr::str_to_title(this_species)` species"
bibliography: "`r find_root_file("references.bib", criterion = has_file("inbomd_examples.Rproj"))`"
---

# Sepal width and length: an explorative analysis

The average sepal width and length are `r round(mean(selection$Sepal.Width),2)` and `r round(mean(selection$Sepal.Width),2)` respectively. Figure \@ref(fig:sepal) shows all `r stringr::str_to_title(this_species)` observations.

```{r sepal, echo = FALSE, warning=FALSE, message=FALSE, fig.height=3, fig.width=5, fig.cap = "Sepal width versus sepal length."}
selection %>%
  ggplot(aes(x = Sepal.Width, y = Sepal.Length)) +
  geom_point()
```

# A simple linear regression model

```{r regression, echo = FALSE, warning=FALSE, message =FALSE}
reg <- lm(data = selection,
   formula = Sepal.Width ~ Sepal.Length + Petal.Length + Petal.Width)
summary_regression <- summary(reg)$coefficients
#we're not interested in the intercept
significant <- which(summary_regression[-1,4] < 0.05)

if (length(significant) == 0) {
   conclusion <- "none of the covariates are significant"
} else if (length(significant) == 1) {
     conclusion <- paste("only", names(significant), "is significant")
} else {
       conclusion <- paste(length(significant), "of the covariates are significant")
       }
summary_regression %>%
  knitr::kable(
    caption = "Regression results.",
    booktabs = TRUE,
    digits = 3) 
```

A simple linear regression model that predicts the sepal width based on sepal length, petal length and petal width shows that `r conclusion` (see Table \@ref(tab:regression)).
