# R based output

## Code

### Chunks + output {.allowframebreaks}

```{r datasets, error = TRUE, message = TRUE}
# logical
c(TRUE, FALSE)
# integer
0:1
# numeric
c(0.0, 1.1)
# scientific
c(1e-10, 1e10)
# character
c("monday", "tuesday", "wednesday")
# factor
factor(c("monday", "tuesday", "wednesday"))
# function
my_fun <- function(x){
  cat("my function is", x)
}
# data.frame
state <- data.frame(
  region = state.region,
  Division = state.division,
  state.x77
)
# function
my_fun("cool")
# messages
message("this is a message")
warning("this is a warning")
stop("this is an error")
# programma flow
if (is.data.frame(state)) {
  summary(state)
} else {
  stop("state is not a data.frame")
}
for (i in 1:5) {
  cat("i =", i, "\n")
}
# number of characters
#       10        20        30        40        50        60        70        80        90       100       110       120
#2345678901234567890123456789012345678901234567890123456789012345678901234567890123456789012345678901234567890123456789012345678901234567890
```

### Using values calculated by R in plain text

```{r echo = FALSE}
life_exp <- sprintf(
  "%.1f (%.1f; %.1f) years",
  mean(state$Life.Exp),
  quantile(state$Life.Exp, 0.025),
  quantile(state$Life.Exp, 0.975)
)
```

Mixing R output and text is straightforward. E.g $1 + 1 = `r 1+1`$. Just make sure that you can use a single and simple command. Precalculate the output in case you need more elaborate calculations. E.g the life expectancy in the USA is `r life_exp`.

## Figures

### Figure with default colour

```{r figtransparant, echo = FALSE, message=FALSE, warning=FALSE}
ggplot(esoph, aes(x = ncases / (ncases + ncontrols))) + geom_histogram() + 
  facet_wrap(~agegp, scales = "free")
```

### Figure with discrete colours

```{r echo = FALSE, message=FALSE}
ggplot(diamonds, aes(x = price, fill = cut)) + geom_histogram()
```

### Figure with discrete colours

```{r echo = FALSE}
ggplot(mtcars, 
       aes(x = wt, y = mpg, colour = factor(cyl), fill = factor(cyl))) +
  geom_point() + geom_smooth(method = "lm")
```

### Figure with single colour gradient

```{r echo = FALSE}
pp <- function(n) {
  df <- expand.grid(
    x = seq_len(n),
    y = seq_len(n)
  )
  df$z <- df$x + df$y
  df
}
ggplot(pp(20), aes(x = x, y = y, fill = z)) +
  geom_tile(width = 0.8, height = 0.8) +
  scale_fill_gradient()
```

### Figure with colour gradient over a central colour

```{r echo = FALSE}
ggplot(pp(20), aes(x = x, y = y, fill = z)) +
  geom_tile(width = 0.8, height = 0.8) +
  scale_fill_gradient2(
    midpoint = 20, 
    low = vl.lightblue, 
    high = vl.lightred
  )
```

### Default order of discrete colours

```{r echo = FALSE, warning=FALSE}
testgrid <- expand.grid(x = 1:4, y = 1:4)
testgrid$z <- factor((testgrid$x - 1) * 4 + testgrid$y)
ggplot(testgrid, aes(x = x, y = y, fill = z, label = z)) +
  geom_tile(width = 0.7, height = 0.7) +
  geom_text() +
  coord_fixed()
```

## Tables

```{r generate-table, echo = FALSE}
# functies om tabellen te genereren
number_table <- function(rows = 20, cols = 10) {
  data.frame(
    matrix(
      rnorm(rows * cols),
      nrow = rows,
      ncol = cols
    ),
    row.names = paste("rij", seq_len(rows))
  )
}
random_word <- function(n_letters = 5){
  paste(
    sample(letters, size = n_letters, replace = TRUE),
    collapse = ""
  )
}
random_sentence <- function(n_words = 10, n_letters = 10){
  paste(
    sapply(
      rpois(n_words, n_letters), 
      FUN = random_word
    ),
    collapse = " "
  )
}
text_table <- function(rows = 20, cols = 10, n_words = 10, n_letters = 5){
  x <- data.frame(
    matrix(
      sapply(
        rpois(rows * cols, lambda = n_words), 
        random_sentence,
        n_letters
      ),
      nrow = rows,
      ncol = cols
    )
  )
  colnames(x) <- head(LETTERS, cols)
  return(x)
}
generate_table <- function(rows = 20, cols = c(5, 5), n_words = 10, n_letters = 5){
  cbind(
    number_table(rows = rows, cols = cols[1]),
    text_table(
      rows = rows, cols = cols[2], n_words = n_words, n_letters = n_letters
    )
  )
}
```

### Default `kable()`

```{r kable, echo = FALSE}
kable(
  generate_table(15, c(2, 1), 4, 4),
  digits = 2
)
```

### Booktabs `kable()`

```{r kable-booktabs, echo = FALSE}
kable(
  generate_table(15, c(2, 1), 4, 4),
  digits = 2,
  booktabs = TRUE
)
```

### longtable `kable()`

```{r kable-longtable, echo = FALSE}
kable(
  generate_table(15, c(2, 1), 4, 4), 
  digits = 2,
  longtable = TRUE
)
```

### Markdown `kable()`

```{r kable-markdown, echo = FALSE}
kable(
  generate_table(15, c(2, 1), 4, 4), 
  digits = 2
)
```

### Pandoc `kable()`

```{r kable-pandoc, echo = FALSE}
kable(
  generate_table(15, c(2, 1), 4, 4), 
  digits = 2,
  format = "pandoc"
)
```
